---
title: "Toronto Police Data"
format: html
---
#### Preamble ####
# Purpose: Read in data from Toronto investigated 
# complaints between 2014 to 2022
# Author: Yi Fei Pang
# Email: yifei.pang@mail.utoronto.ca
# Date: 15 January 2023
# Prerequisites: Have access to the opendatatoronto dataset


#### Workspace Setup ####
install.packages("tidyverse")\
install.packages("janitor")\
install.packages("opendatatoronto")\
install.packages("dplyr")\
install.packages("ggplot2")


#### Read in Data ####
```{r}
library(opendatatoronto)
library(dplyr)
library(tidyverse)
library(janitor)
library(ggplot2)

package <- show_package("police-annual-statistical-report-total-public-complaints")

# get all resources for this package
resources <- list_package_resources("police-annual-statistical-report-total-public-complaints")

# identify datastore resources; by default, Toronto Open Data sets datastore 
# resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the first datastore resource
data <- filter(datastore_resources, row_number()==1) %>% get_resource()
```


#### Clean Data ####
```{r}
cleaned_police_data <-
  data |>
  select(
    YEAR,
    TYPE,
    COUNT_
  ) %>%
  # must mutate because the COUNT_ is <chr> data type and not <int>
  mutate(COUNT_ = as.numeric(COUNT_)) %>%
  group_by(YEAR, TYPE) %>%
  summarise(CASES = sum(COUNT_))

```

#### Display Data ####
```{r}
ggplot(cleaned_police_data, aes(x = factor(YEAR), y = CASES, fill = TYPE)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Cases By Year",
       x = "Year",
       y = "Cases",
       fill = "Complaint Type") +
  theme_minimal()

```

